Designing and modeling a browser-based distributed evolutionary computation systems

### 2015-04-05 modeling-volunteer-computing-gecco.tex
Finishing server description for checking by @fergunet

All in all, it amounts to a few hundred lines of code
including all languages (client and server JavaScript, HTML and CSS). This code can
be deployed, in principle, in any server that is able to run node.js, but for the
purposes of this experiment OpenShift (\url{http://openshift.com}), a Platform as a Service with a
free tier, has been chosen. It could be also be deployed, with small
or no modification, to Heroku or other similar PaaS. Any of them are
free up to a certain level of use and thus fulfill the requirement of
using free resources as a sustainable way of performing distributed
evolutionary compuation experiments \cite{Merelo2014,jj:idc:lowcost}.

Having the pages and the server in the same domain allows to work in the
default Ajax mode, that does not allow cross-site requests. However,
these can also be enabled if needed so that the static pages and the
REST server can be hosted in different servers. In this case, every domain holds
a single experiment, there is no multi-tenancy of experiments, so its
management is quite simple.

During and after the experiment, the data can be collected from one of
the logs: the log file (if the PaaS allows writing to the filesystem,
which is not the case in Heroku) or to online logging services such as
Loggly (\url{http://loggly.com}, which is the one we are using. Loggly was chosen for its
easiness and support from the {\tt winston} logging library; however,
it is only free for a month, so a more sustainable alternative (read:
free) will have to be sought in the future. Other online logging
services have a free tier that can be used for ever; however, its
support and facilities are not so good. Choosing a good logging
service for experiments in open science is an open issue but it is
outside the scope of this paper. It will have to be left for future
work. 
% Include loggly graphs here.
As indicated in the introduction, every part of the experiment and
data gathered have been released as free software in the GitHub
repository \url{https://github.com/JJ/splash-volunteer}. 
will be analyzed in the next section.

>>>>

All in all, it amounts to a few hundred lines of code
including all languages (client and server JavaScript, HTML and CSS). This code can
be deployed, in principle, in any server that is able to run node.js.

In this experiment we have used OpenShift (\url{http://openshift.com})
and Heroku (\url{http://heroku.com}) which are Platform as a Service products with a
free tier. It could be also be deployed, with small
or no modification, to other similar PaaS, maybe with a bit of adaptation. Any of them are
free up to a certain level of use and thus fulfill the requirement of
using free resources as a sustainable way of performing distributed
evolutionary compuation experiments \cite{Merelo2014,jj:idc:lowcost}.

Programming the server so that it can be deployed, without
modification, to any PaaS is not difficult, but you have to consider
the different capabilities of each one.\begin{itemize}
\item Heroku does not have a permanent filesystem, so if you want to
  create any directory it has to be done when the server is
  started. It is only guaranteed to last while the application is
  running. 
\item In order to scale properly, all requests in Openshift are routed
  through a proxy. This means that the requests all have, as remote
  address, this IP. Other ways of obtaining the requests must be found
  out. On the other hand, it does have a permanent directory for data.
\end{itemize}

Every PaaS has its own way of defining environment variables, which is the way
used for passing private information such as passwords to the
server. This can be done usually through a command-line client ({\tt
  heroku} or {\tt rhc}), which is also used to check the logs or
restart the client. Eventually, we decided to leave the two servers
running.

Having the pages and the server in the same domain allows to work in the
default Ajax mode, that does not allow cross-site requests. However,
these can also be enabled if needed so that the static pages and the
REST server can be hosted in different servers, allowing the genetic
algorithm code to {\em parasite} other pages.

In this case, every domain holds
a single experiment, there is no multi-tenancy of experiments, so its
management is quite simple. A different experiment would need to
deploy and start a server that serves different code and checks for
different termination conditions.

During and after the experiment, the data can be collected from one of
the logs: the server log file, data files (that cannot be used in
Heroku, for instance)  or from online logging services such as
Loggly (\url{http://loggly.com}), which is the one we are using. Loggly was chosen for its
easiness and support from the {\tt winston} logging library; however,
it is only free for a month, so a more sustainable alternative (read:
free) will have to be sought in the future, after this paper is finished. Other online logging
services have a free tier that can be used for ever; however, its
support and graphing facilities were not so complete. Choosing a good logging
service for experiments in open science is, for us, an open issue but it is
outside the scope of this paper. It will have to be left for future
work. 
% Include loggly graphs here.
As indicated in the introduction, every part of the experiment and
data gathered have been released as free software in the GitHub
repository \url{https://github.com/JJ/splash-volunteer}. 
will be analyzed in the next section.
